Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-09 vesicle_code/docs/permeableVes/main.tex
updating figure a little bit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
Based on our analysis, a semi-permeable vesicle in a quiescent flow
relaxes to a circle with constant tension $\Lambda = 2\pi^2/L^2$.
Furthermore, equation~\eqref{eqn:areaROC} shows that the area change is
proportional to $\beta$, so we expect the times for the vesicle to
deviate from its initial reduced area and to reach its steady state to
both be proportional to $\beta^{-1}$. This is verified by considering a
vesicle with initial reduced area $\nu = 0.65$ and several different
semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
the time, $t_\mathrm{steady}$, required for the vesicle's area to be
within 1.3\% of its steady state reduced area, and the time,
$t_\mathrm{inflate}$, required for the vesicle's reduced area to exceed
1.3\% of its initial reduced area. Both times are proportional to
$\beta^{-1}$. 
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

In Figure~\ref{fig:relaxationComposite}, we examine the transient
dynamics of four vesicles $\beta = 10^{-3}$ and a variety of initial
configurations.  First, in all cases the vesicle eventually reaches a
circular shape of reduced area $\nu=1$. Second, we observe that even
though the vesicle is inflating to a circle, meaning that there is a net
flux into the vesicle, there are regions where the flux is out of the
vesicle. These regions are denoted in blue and are only present at early
times.  Finally, the vesicles area does not initially change by much,
and intermediate shapes of semi-permeable vesicles resemble the shape of
an impermeable vesicle. This can be explained by considering the two
terms that contribute to the vesicle velocity: (i) the permeability
velocity, $\beta(\ff \cdot \nn) \nn$, and (ii) the force balance
velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$, the permeability velocity
is much smaller than the force balance velocity at early times, and the
semi-permeable vesicle dynamics resembles those of an impermeable
vesicle. As the force balance velocity decreases, the permeability
velocity dominates, and the vesicle begins to inflate to a circle with
no vanishing force and velocity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure}[htp]
  \ifTikz
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady-state inclination
  angle for four different flow velocities. The steady-state shapes are
  superimposed. (c) The reduced area of the slipper formed with the
  flow rate of 800 $\mu m/s$. (d) The vesicle shape shown with the
  background imposed flow. The channel width is 12.5 times larger than
  the vesicle radius. The corresponding reduced area for each shape is
  indicated by the marks in plot (c).}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
Our analysis indicates that a semi-permeable vesicle in a quiescent flow
relaxes to a circle with constant tension $\Lambda = 2\pi^2/L^2$.
Furthermore, equation~\eqref{eqn:areaROC} shows that the area change is
proportional to $\beta$, and the times for the vesicle to deviate from
its initial reduced area and to reach its steady state should both be
proportional to $\beta^{-1}$. This is verified for a vesicle with
initial reduced area $\nu = 0.65$ and with different semi-permeability
rates. In Table~\ref{tbl:ellipseRelaxTimes} we report the required time,
$t_\mathrm{steady}$, for the vesicle's reduced area to be within 1.3\%
of its steady state reduced area, and the required time,
$t_\mathrm{inflate}$, for the vesicle's reduced area to exceed 1.3\% of
its initial reduced area. Both times are proportional to $\beta^{-1}$,
and for this reason, our numerical results are reported in terms of the
scaled time $\beta t$.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

In Figure~\ref{fig:relaxationComposite}, we examine the transient
dynamics of four vesicles with $\beta = 10^{-3}$ and a variety of
initial configurations. As expected, each vesicle eventually
reaches a circular shape with reduced area $\nu=1$. Next, we observe
that even though the vesicle is inflating to a circle, meaning that
there is a net flux into the vesicle, there are regions where the flux
is out of the vesicle. These regions are denoted in blue and are only
present at early times. Finally, the vesicles's reduced area do not 
change by much at early times, and intermediate shapes of semi-permeable vesicles
resemble the shape of an impermeable vesicle. This can be explained by
considering the two terms that contribute to the vesicle velocity: (i)
the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and (ii) the
force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$, the
permeability velocity is much smaller than the force balance velocity at
early times, and the semi-permeable vesicle dynamics resembles those of
an impermeable vesicle. As the force balance velocity decreases, the
permeability velocity dominates, and the vesicle begins to inflate to a
circle with no vanishing force and velocity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
%  \ifTikz
  \input{figures/parabolicComposite.tikz}
%  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady-state inclination
  angle for four different flow velocities. The steady-state shapes are
  superimposed. (c) The reduced area of the slipper formed with the
  flow rate of 800 $\mu m/s$. (d) The vesicle shape shown with the
  background imposed flow. The channel width is 12.5 times larger than
  the vesicle radius. The corresponding reduced area for each shape is
  indicated by the marks in plot (c).}
  \todo[inline]{shift right plot to the right and increase necessary
  font sizes}
\end{figure*}
